---
output: 
  github_document:
    toc: FALSE
---

<!-- badges: start -->
  [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
  <!-- badges: end -->  

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  eval = T
)

library(tidyverse)
```



# Part 0. Proposal

Proposing the {sf2stat} package! 🦄 
<!-- (typical package introduction write up; but actually aspirational) -->

The goal of {sf2stat} is to make it easier to prep *sf data* for use in a ggproto Stat computation; the Stat then can be used for creating a stat/geom function to be used in ggplot2 plots.

Without the package, we live in the effortful world, in which we'd have to prep our own data including figuring out the bounding box for each geometry, and, if we want labeling functionality, the centroid for each geometry.  

With the {sf2stat} package, we'll live in a different world (🦄 🦄 🦄) where the task is a snap 🫰: 

Proposed API:

```

library(sf2stat)

my_geom_ref_data <- sf_df_prep_for_stat(data, id_col_name = county_name)

```



# Part I. Work out functionality  ✅ 

In this section we'll use the nc sf dataframe to check out how our functions work.

```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package="sf")) %>%
  select(NAME, FIPS)

nc
```


First we have a function that takes an sf data frame and adds columns x and y for the centroids of the geometries.

```{r sf_df_add_xy_center_coords}
sf_df_add_xy_center_coords <- function(sf_df){

sf_df |>
    dplyr::pull(geometry) |>
    sf::st_zm() |>
    sf::st_point_on_surface() ->
  points_sf

the_coords <- do.call(rbind, sf::st_geometry(points_sf)) |>
  tibble::as_tibble() |> setNames(c("x","y"))

cbind(sf_df, the_coords)

}
```


```{r}
nc |> sf_df_add_xy_center_coords()
```


Second we have a function that's going to return bounding boxes as a dataframe.  For our reference data we need these xmin, xmax variables for each row in our data.

```{r sf_df_return_bbox_df}
sf_df_return_bbox_df <- function(sf_df){
  
  bb <- sf::st_bbox(sf_df)

  data.frame(xmin = bb[1], ymin = bb[2],
             xmax = bb[3], ymax = bb[4])

}
```

```{r}
nc[10,] |> sf_df_return_bbox_df()
```



```{r sf_df_prep_for_stat}
sf_df_prep_for_stat <- function(sf_df, id_col_name = NULL){
  
  sf_df |>
    # using purrr allows us to get bb for each row
    dplyr::mutate(bb =
                    purrr::map(geometry,
                               sf_df_return_bbox_df)) |>
    tidyr::unnest(bb) |>
    data.frame() |>
    sf_df_add_xy_center_coords() ->
  sf_df_w_bb_and_centers

  # use first column as keep/drop column unless otherwise specified
  if(is.null(id_col_name)){id_col_name <- 1} 
  
  sf_df_w_bb_and_centers$id_col <- sf_df_w_bb_and_centers[,id_col_name]

  return(sf_df_w_bb_and_centers)
  
  }
```


```{r template_compute_panel_code}
template_compute_panel_code <- function(){
  
"compute_panel_geo_XXXX <- function(data, scales, keep_id = NULL, drop_id = NULL){
  
  if(!is.null(keep_id)){ data <- filter(data, id_col %in% keep_id) }
  if(!is.null(drop_id)){ data <- filter(data, !(id_col %in% drop_id)) }
  
  if(!stamp){data <- dplyr::inner_join(data, geo_ref_XXXX)}
  if( stamp){data <- geo_ref_XXXX }
  
  data
  
}" |> cat()
  
}
```

```{r template_stat_code}
template_stat_code <- function(){
  
'StatXXXXsf <- ggplot2::ggproto(`_class` = "StatXXXXsf",
                                `_inherit` = ggplot2::Stat,
                                required_aes = c("fips|county_name|XXXX"),
                                compute_panel = compute_panel_geo_XXXX,
                               default_aes = c(label = ggplot2::after_stat(id_col)))' |> cat()
}
```

```{r template_layer_code}
template_layer_code <- function(){ 'stat_XXXX <- function(
      mapping = NULL,
      data = NULL,
      geom = ggplot2::GeomSf,
      position = "identity",
      na.rm = FALSE,
      show.legend = NA,
      inherit.aes = TRUE,
      crs = "NAD27", # "NAD27", 5070, "WGS84", "NAD83", 4326 , 3857
      ...) {

  c(ggplot2::layer_sf(
              stat = StatXXXX,  # proto object from step 2
              geom = geom,  # inherit other behavior
              data = data,
              mapping = mapping,
              position = position,
              show.legend = show.legend,
              inherit.aes = inherit.aes,
              params = rlang::list2(na.rm = na.rm, ...)
              ),
              
              coord_sf(crs = crs,
                       default_crs = sf::st_crs(crs),
                       datum = crs,
                       default = TRUE)
     )
  }' |> cat()

}
  
```



## Try it out: how to use in gg* package

Let's see how we might recreate the functionality in the ggnorthcarolina package


### Step 00. prep reference data

```{r, eval = F}
usethis::use_data_raw()
```

```{r nc_geo_reference_prep, eval = T}
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))

nc |>
  dplyr::select(county_name = NAME, fips = FIPS) |>
  sf_df_prep_for_stat(id_col_name = "county_name") ->
nc_geo_reference
```


```{r, eval = F}
usethis::use_data(nc_geo_reference)
```

# Step 1 and 2

```{r StatNcsf}
# step 1
compute_panel_nc <- function(data, scales, keep_id = NULL, drop_id = NULL, stamp = FALSE){
  
  if(!stamp){data <- dplyr::inner_join(data, nc_geo_reference)}
  if( stamp){data <- nc_geo_reference }
  
  if(!is.null(keep_id)){ data <- filter(data, id_col %in% keep_id) }
  if(!is.null(drop_id)){ data <- filter(data, !(id_col %in% drop_id)) }
  
  data
  
}

# step 2
StatNcsf <- ggplot2::ggproto(`_class` = "StatNcsf",
                                `_inherit` = ggplot2::Stat,
                                # required_aes = c("fips|county_name"),
                                compute_panel = compute_panel_nc,
                               default_aes = ggplot2::aes(label = after_stat(id_col)))
```



## Step 3

```{r, eval = F}
readme2pkg::chunk_to_r("stat_county")
```

```{r stat_county}
# step 3
stat_county <- function(
      mapping = NULL,
      data = NULL,
      geom = ggplot2::GeomSf,
      position = "identity",
      na.rm = FALSE,
      show.legend = NA,
      inherit.aes = TRUE,
      crs = "NAD27", # "NAD27", 5070, "WGS84", "NAD83", 4326 , 3857
      ...) {

  c(ggplot2::layer_sf(
              stat = StatNcsf,  # proto object from step 2
              geom = geom,  # inherit other behavior
              data = data,
              mapping = mapping,
              position = position,
              show.legend = show.legend,
              inherit.aes = inherit.aes,
              params = rlang::list2(na.rm = na.rm, ...)
              ),
              
              coord_sf(crs = crs,
                       default_crs = sf::st_crs(crs),
                       datum = crs,
                       default = TRUE)
     )
  }
```



# test it out

```{r}
library(ggplot2)
nc |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(fips = FIPS) +
  stat_county() + 
  aes(fill = BIR79)
```


### Make derivitive functions, aliases


```{r geom_county}
geom_county <- stat_county
geom_county_label <- function(...){stat_county(geom = "text",...)}
stamp_county <- function(...){
  stat_county(stamp = T, 
              data = mtcars,
              aes(fill = NULL, color = NULL, label = NULL, 
                  fips = NULL, county_name = NULL), 
              ...)}
stamp_county_label <- function(...){
  stat_county(stamp = T, 
              geom = "text", 
              data = mtcars, 
              aes(fill = NULL, color = NULL,
                  fips = NULL, county_name = NULL), 
              ...)}
```



# try those out

```{r}
nc |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(fips = FIPS) +
  geom_county() + 
  geom_county_label(check_overlap = T,
                    color = "grey85") +
  aes(fill = BIR79) 

last_plot() + 
  stamp_county() + 
  stamp_county_label()

ggplot() + 
  stamp_county()

last_plot() + 
  stamp_county_label(check_overlap = T)

last_plot() + 
  stamp_county(keep_id = "Wake", fill = "darkred")
```

# Wanting even more?

## Stamps for each polygon?

```{r stamp_county_wake}
#' Title
#'
#' @param ... 
#'
#' @return
#' @export
#'
#' @examples
stamp_county_wake <- function(...){stamp_county(keep_id = 'Wake', ...)}

#' Title
#'
#' @param ... 
#'
#' @return
#' @export
#'
#' @examples
stamp_county_label_wake <- function(...){stamp_county_label(keep_id = 'Wake', ...)}

```



```{r eval=F}
nc_geo_reference$county_name 


readme2pkg::chunk_variants_to_dir("stamp_county_wake", 
                              replace1 = "wake",
                              replacements1 =
                                str_replace(
                                  tolower(nc_geo_reference$county_name), 
                                  " ", "_"),
                              replace2 = "Wake", 
                              replacements2 = nc_geo_reference$county_name)
```



# Part II. Packaging and documentation  🚧 ✅ 

## Phase 1. Minimal working package

### Bit A. Created package archetecture, running `devtools::create(".")` in interactive session. 🚧 ✅


```{r, eval = F}
devtools::create(".")
```

### Bit B. Added roxygen skeleton? 🚧 ✅

Use a roxygen skeleton for auto documentation and making sure proposed functions are *exported*.  Generally, early on, I don't do much (anything) in terms of filling in the skeleton for documentation, because things may change.

### Bit C. Managed dependencies ? 🚧 ✅

Package dependencies managed, i.e. `depend::function()` in proposed functions and declared in the DESCRIPTION

```{r, eval = F}
usethis::use_package("sf")
usethis::use_package("dplyr")
usethis::use_package("tibble")
usethis::use_package("tidyr")
usethis::use_package("purrr")

```


### Bit D. Moved functions R folder? 🚧 ✅

Use new {readme2pkg} function to do this from readme...

```{r, eval = F}
readme2pkg::chunk_to_r("sf_df_return_bbox_df")
readme2pkg::chunk_to_r("sf_df_add_xy_center_coords")
readme2pkg::chunk_to_r("sf_df_prep_for_stat")
readme2pkg::chunk_to_r("template_compute_panel_code")
readme2pkg::chunk_to_r("template_stat_code")
readme2pkg::chunk_to_r("template_layer_code")
```



### Bit E. Run `devtools::check()` and addressed errors. 🚧 ✅

```{r, eval = F}
devtools::check(pkg = ".")
```

### Bit F. Build package 🚧 ✅

```{r, eval = F}
devtools::build()
```


### Bit G. Write traditional README that uses built package (also serves as a test of build. 🚧 ✅ 

The goal of the {xxxx} package is to ...

Install package with:

```
remotes::installgithub("EvaMaeRey/readme2pkg.template")
```

Once functions are exported you can remove go to two colons, and when things are are really finalized, then go without colons (and rearrange your readme...)

```{r, eval = F}
library(sf2stat)  ##<< change to your package name here

nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))

nc |>
  dplyr::select(county_name = NAME, fips = FIPS) |>
  sf2stat::sf_df_prep_for_stat(id_col_name = "county_name") ->
nc_geo_reference
```




### Bit H. Chosen a license? 🚧 ✅


```{r, eval = F}
usethis::use_mit_license()
```

### Bit I. Add lifecycle badge (experimental)

```{r, eval = F}
usethis::use_lifecycle_badge("experimental")
```


## Phase 2: Listen & iterate 🚧 ✅

Try to get feedback from experts on API, implementation, default decisions.  Is there already work that solves this problem? 


## Phase 3: Let things settle

### Bit A. Settle on examples.  Put them in the roxygen skeleton and readme. 🚧 ✅

### Bit B. Written formal tests of functions and save to test that folders 🚧 ✅


That would look like this...

```{r test_calc_times_two_works, eval = F}
library(testthat)

test_that("calc times 2 works", {
  expect_equal(times_two(4), 8)
  expect_equal(times_two(5), 10)
  
})
```


```{r, eval = F}
readme2pkg::chunk_to_tests_testthat("test_calc_times_two_works")
```


### Bit C. Added a description and author information in the DESCRIPTION file 🚧 ✅

### Bit D. Addressed *all* notes, warnings and errors. 🚧 ✅

## Phase 4. Promote to wider audience...

### Bit A. Package website built? 🚧 ✅

### Bit B. Package website deployed? 🚧 ✅

## Phase 5: Harden/commit

### Submit to CRAN/RUniverse? 🚧 ✅

# Appendix: Reports, Environment

## Edit Description file

```{r, eval = F}
readLines("DESCRIPTION")
```

## Environment

Here I just want to print the packages and the versions

```{r}
all <- sessionInfo() |> print() |> capture.output()
all[11:17]
```

## `devtools::check()` report

```{r, error = T, results="hide", warning=F, eval=F}
devtools::check(pkg = ".")
```





