---
output: 
  github_document:
    toc: FALSE
---

<!-- badges: start -->
  [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
  <!-- badges: end -->  

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  eval = T
)

library(tidyverse)
```

# *To the reader* 

Welcome to the R package building helper *readme2pkg.template*!

Below, is a readme that provides steps for building a package.  This readme acts as a worksheet, checklist, and control document as functions used in package building are included within.  

We'll use the `{readme2pkg}` helper package to send code chunks to different directories in the package.  

To install `{readme2pkg}`:

```

remotes::install.github("EvaMaeRey/readme2pkg")

```

# Part 0. Proposal

Proposing the {sf2stat} package! 🦄 
<!-- (typical package introduction write up; but actually aspirational) -->

The goal of {sf2stat} is to make it easier to prep sf data to become part of a ggproto Stat object, which in turn can be used for creating a stat/geom function.


Without the package, we live in the effort-ful world, in which we'd have to prep our own data including figuring out the bounding box for each geometry, and the centroid for each geometry.  


  
With the {sf2stat} package, we'll live in a different world (🦄 🦄 🦄) where the task is a snap 🫰: 

Proposed API:

```

library(sf2stat)

my_geom_ref_data <- sf_df_prep_for_stat(data, id_col_name = county_name)

```



# Part I. Work out functionality  ✅ 

Here is a function that will do some work...


```{r sf_df_return_bbox_df}
sf_df_return_bbox_df <- function(sf_df){
  
  bb <- sf::st_bbox(sf_df)

  data.frame(xmin = bb[1], ymin = bb[2],
             xmax = bb[3], ymax = bb[4])

}
```


```{r sf_df_add_xy_center_coords}
sf_df_add_xy_center_coords <- function(sf_df){

sf_df |>
    dplyr::pull(geometry) |>
    sf::st_zm() |>
    sf::st_point_on_surface() ->
  points_sf

the_coords <- do.call(rbind, sf::st_geometry(points_sf)) |>
  tibble::as_tibble() |> setNames(c("x","y"))

cbind(sf_df, the_coords)

}
```

```{r sf_df_prep_for_stat}
sf_df_prep_for_stat <- function(sf_df, id_col_name = NULL){
  
  sf_df |>
    # using purrr allows us to get bb for each row
    dplyr::mutate(bb =
                    purrr::map(geometry,
                               sf_df_return_bbox_df)) |>
    tidyr::unnest(bb) |>
    data.frame() |>
    sf_df_add_xy_center_coords() ->
  sf_df_w_bb_and_centers

  # use first column as keep/drop column unless otherwise specified
  if(is.null(id_col_name)){id_col_name <- 1} 
  
  sf_df_w_bb_and_centers$id_col <- sf_df_w_bb_and_centers[,id_col_name]

  return(sf_df_w_bb_and_centers)
  
  }
```


```{r template_compute_panel_code}
template_compute_panel_code <- function(){
  
"compute_panel_geo_XXXX <- function(data, scales, keep_id = NULL, drop_id = NULL){
  
  if(!is.null(keep_id)){ data <- filter(data, id_col %in% keep_id) }
  if(!is.null(drop_id)){ data <- filter(data, !(id_col %in% drop_id)) }
  
  if(!stamp){data <- dplyr::inner_join(data, geo_ref_XXXX)}
  if( stamp){data <- geo_ref_XXXX }
  
  data
  
}" |> cat()
  
}
```

```{r template_stat_code}
template_stat_code <- function(){
  
'StatXXXXsf <- ggplot2::ggproto(`_class` = "StatXXXXsf",
                                `_inherit` = ggplot2::Stat,
                                required_aes = c("fips|county_name|XXXX"),
                                compute_panel = compute_panel_geo_XXXX,
                               default_aes = c(label = ggplot2::after_stat(id_col)))' |> cat()
}
```

```{r template_layer_code}
template_layer_code <- function(){ 'stat_XXXX <- function(
      mapping = NULL,
      data = NULL,
      geom = ggplot2::GeomSf,
      position = "identity",
      na.rm = FALSE,
      show.legend = NA,
      inherit.aes = TRUE,
      crs = "NAD27", # "NAD27", 5070, "WGS84", "NAD83", 4326 , 3857
      ...) {

  c(ggplot2::layer_sf(
              stat = StatXXXX,  # proto object from step 2
              geom = geom,  # inherit other behavior
              data = data,
              mapping = mapping,
              position = position,
              show.legend = show.legend,
              inherit.aes = inherit.aes,
              params = rlang::list2(na.rm = na.rm, ...)
              ),
              
              coord_sf(crs = crs,
                       default_crs = sf::st_crs(crs),
                       datum = crs,
                       default = TRUE)
     )
  }' |> cat()

}
  
```



## Try it out

```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))

nc |>
  dplyr::select(county_name = NAME, fips = FIPS) |>
  sf_df_prep_for_stat(id_col_name = "county_name") ->
nc_geo_reference

# step 1
compute_panel_nc <- function(data, scales, keep_id = NULL, drop_id = NULL, stamp = FALSE){
  
  if(!stamp){data <- dplyr::inner_join(data, nc_geo_reference)}
  if( stamp){data <- nc_geo_reference }
  
  if(!is.null(keep_id)){ data <- filter(data, id_col %in% keep_id) }
  if(!is.null(drop_id)){ data <- filter(data, !(id_col %in% drop_id)) }
  
  data
  
}

# step 2
StatNcsf <- ggplot2::ggproto(`_class` = "StatNcsf",
                                `_inherit` = ggplot2::Stat,
                                # required_aes = c("fips|county_name"),
                                compute_panel = compute_panel_nc,
                               default_aes = ggplot2::aes(label = after_stat(id_col)))

# step 3
stat_county <- function(
      mapping = NULL,
      data = NULL,
      geom = ggplot2::GeomSf,
      position = "identity",
      na.rm = FALSE,
      show.legend = NA,
      inherit.aes = TRUE,
      crs = "NAD27", # "NAD27", 5070, "WGS84", "NAD83", 4326 , 3857
      ...) {

  c(ggplot2::layer_sf(
              stat = StatNcsf,  # proto object from step 2
              geom = geom,  # inherit other behavior
              data = data,
              mapping = mapping,
              position = position,
              show.legend = show.legend,
              inherit.aes = inherit.aes,
              params = rlang::list2(na.rm = na.rm, ...)
              ),
              
              coord_sf(crs = crs,
                       default_crs = sf::st_crs(crs),
                       datum = crs,
                       default = TRUE)
     )
  }


library(ggplot2)
nc |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(fips = FIPS) +
  stat_county() + 
  aes(fill = BIR79)


geom_county <- stat_county
geom_county_label <- function(...){stat_county(geom = "text",...)}
stamp_county <- function(...){stat_county(data = mtcars, stamp = T, ...)}
stamp_county_label <- function(...){stat_county(geom = "text", data = mtcars, stamp = T, ...)}


nc |>
  sf::st_drop_geometry() |>
  ggplot() +
  aes(fips = FIPS) +
  geom_county() + 
  geom_county_label(check_overlap = T,
                    color = "grey85") +
  aes(fill = BIR79) 

ggplot() + 
  stamp_county()

last_plot() + 
  stamp_county_label(check_overlap = T)

last_plot() + 
  stamp_county(keep_id = "Wake", fill = "darkred")

```


# Part II. Packaging and documentation  🚧 ✅ 

## Phase 1. Minimal working package

### Bit A. Created package archetecture, running `devtools::create(".")` in interactive session. 🚧 ✅


```{r, eval = F}
devtools::create(".")
```

### Bit B. Added roxygen skeleton? 🚧 ✅

Use a roxygen skeleton for auto documentation and making sure proposed functions are *exported*.  Generally, early on, I don't do much (anything) in terms of filling in the skeleton for documentation, because things may change.

### Bit C. Managed dependencies ? 🚧 ✅

Package dependencies managed, i.e. `depend::function()` in proposed functions and declared in the DESCRIPTION

```{r, eval = F}
usethis::use_package("sf")
usethis::use_package("dplyr")
usethis::use_package("tibble")
usethis::use_package("tidyr")
usethis::use_package("purrr")

```


### Bit D. Moved functions R folder? 🚧 ✅

Use new {readme2pkg} function to do this from readme...

```{r, eval = F}
readme2pkg::chunk_to_r("sf_df_return_bbox_df")
readme2pkg::chunk_to_r("sf_df_add_xy_center_coords")
readme2pkg::chunk_to_r("sf_df_prep_for_stat")
readme2pkg::chunk_to_r("template_compute_panel_code")
readme2pkg::chunk_to_r("template_stat_code")
readme2pkg::chunk_to_r("template_layer_code")
```



### Bit E. Run `devtools::check()` and addressed errors. 🚧 ✅

```{r, eval = F}
devtools::check(pkg = ".")
```

### Bit F. Build package 🚧 ✅

```{r, eval = F}
devtools::build()
```


### Bit G. Write traditional README that uses built package (also serves as a test of build. 🚧 ✅ 

The goal of the {xxxx} package is to ...

Install package with:

```
remotes::installgithub("EvaMaeRey/readme2pkg.template")
```

Once functions are exported you can remove go to two colons, and when things are are really finalized, then go without colons (and rearrange your readme...)

```{r, eval = F}
library(sf2stat)  ##<< change to your package name here

nc <- sf::st_read(system.file("shape/nc.shp", package="sf"))

nc |>
  dplyr::select(county_name = NAME, fips = FIPS) |>
  sf2stat::sf_df_prep_for_stat(id_col_name = "county_name") ->
nc_geo_reference
```




### Bit H. Chosen a license? 🚧 ✅


```{r, eval = F}
usethis::use_mit_license()
```

### Bit I. Add lifecycle badge (experimental)

```{r, eval = F}
usethis::use_lifecycle_badge("experimental")
```


## Phase 2: Listen & iterate 🚧 ✅

Try to get feedback from experts on API, implementation, default decisions.  Is there already work that solves this problem? 


## Phase 3: Let things settle

### Bit A. Settle on examples.  Put them in the roxygen skeleton and readme. 🚧 ✅

### Bit B. Written formal tests of functions and save to test that folders 🚧 ✅


That would look like this...

```{r test_calc_times_two_works, eval = F}
library(testthat)

test_that("calc times 2 works", {
  expect_equal(times_two(4), 8)
  expect_equal(times_two(5), 10)
  
})
```


```{r, eval = F}
readme2pkg::chunk_to_tests_testthat("test_calc_times_two_works")
```


### Bit C. Added a description and author information in the DESCRIPTION file 🚧 ✅

### Bit D. Addressed *all* notes, warnings and errors. 🚧 ✅

## Phase 4. Promote to wider audience...

### Bit A. Package website built? 🚧 ✅

### Bit B. Package website deployed? 🚧 ✅

## Phase 5: Harden/commit

### Submit to CRAN/RUniverse? 🚧 ✅

# Appendix: Reports, Environment

## Edit Description file

```{r, eval = F}
readLines("DESCRIPTION")
```

## Environment

Here I just want to print the packages and the versions

```{r}
all <- sessionInfo() |> print() |> capture.output()
all[11:17]
```

## `devtools::check()` report

```{r, error = T, results="hide", warning=F}
devtools::check(pkg = ".")
```





